---
title: "IDS 702 HW 4"
author: "Your Name Here"
format: pdf
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute.

**Add your name in the Author section in the header**

## Load data

```{r, message=FALSE}
library(tidyverse)
library(modelsummary)
library(caret)
library(pROC)

nba <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/nba_games_stats.csv")
```

## Exercise 1

```{r}
## create subset
nba_cho <- 
  
## create factor variable for Home
  
## create new Win variable (include quality control check)
  
## format date variable
nba_cho <- nba_cho |>
  mutate(Date_clean = as.Date(Date, "%Y-%m-%d"))
```

## Exercise 2

```{r}
## code to fill in the table
```

| Variable                          | Wins (N= 153) | Losses (N= \_\_) |
|-----------------------------------|---------------|------------------|
| Home games - N (%)                | 92 (60)       |                  |
| Team Points - mean (SD)           | 109 (11.4)    |                  |
| Field Goal Percentage - mean (SD) |               |                  |
| Assists - mean (SD)               |               |                  |
| Steals - mean (SD)                |               |                  |
| Blocks - mean (SD)                |               |                  |
| Opponent Points - mean (SD)       |               |                  |
| Total Rebounds - mean (SD)        |               |                  |
| Turnovers - mean (SD)             |               |                  |

## Exercise 3

## Exercise 4

```{r}
## model here
nba_mod1 <-
  
modelsummary(nba_mod1,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)
```

## Exercise 5

```{r}

```

## Exercise 6

## Exercise 7

```{r}

```

## Exercise 8

```{r}

```

## Exercise 9
