---
title: "IDS 702 HW 4 - KEY"
author: "Your Name Here"
format: pdf
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute.

**Add your name in the Author section in the header**

## Load data

```{r, message=FALSE}
library(tidyverse)
library(modelsummary)
library(caret)
library(pROC)

nba <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/nba_games_stats.csv")
```

## Exercise 1

```{r}
## create subset
nba_cho <- nba |> filter(Team=="CHO")
  
## create factor variable for Home
nba_cho$Home <- factor(nba_cho$Home)
  
## create new Win variable (include quality control check)
nba_cho$Win <- ifelse(nba_cho$WINorLOSS=="W",1,0)
nba_cho$Win <- factor(nba_cho$Win, 
                      levels = c(0,1),
                      labels = c("Loss","Win"))
table(nba_cho$WINorLOSS, nba_cho$Win)
  
## format date variable
nba_cho <- nba_cho |>
  mutate(Date_clean = as.Date(Date, "%Y-%m-%d"))
```

## Exercise 2

```{r}
## code to fill in the table
nba_cho |> count(Win)

nba_cho |> 
  group_by(Win) |> 
  count(Home) |> 
  mutate(perc=n/sum(n)*100)

nba_cho |>
  group_by(Win) |>
  summarise(mean(TeamPoints), sd(TeamPoints),
            mean(FieldGoals.), sd(FieldGoals.),
            mean(Assists), sd(Assists),
            mean(Steals), sd(Steals),
            mean(Blocks), sd(Blocks),
            mean(OpponentPoints), sd(OpponentPoints),
            mean(TotalRebounds), sd(TotalRebounds),
            mean(Turnovers), sd(Turnovers))
```

| Variable                          | Wins (N= 153) | Losses (N= \_\_) |
|-----------------------------------|---------------|------------------|
| Home games  - N (%)               | 92 (60)       |                  |
| Team Points - mean (SD)           | 109 (11.4)    |                  |
| Field Goal Percentage - mean (SD) |               |                  |
| Assists - mean (SD)               |               |                  |
| Steals - mean (SD)                |               |                  |
| Blocks - mean (SD)                |               |                  |
| Opponent Points - mean (SD)       |               |                  |
| Total Rebounds - mean (SD)        |               |                  |
| Turnovers - mean (SD)             |               |                  |

## Exercise 3

Similar statistics likely to be highly correlated which could lead to issues with multicollinearity. E.g., Opp.X3PointShots, Opp.X3PointShotsAttempted, Opp.X3PointShots.

## Exercise 4

```{r}
## model here
nba_mod1 <- glm(Win ~ Home+TeamPoints+FieldGoals.+
                  Assists+Steals+Blocks+TotalRebounds+Turnovers,
                data=nba_cho, family="binomial")
  
modelsummary(nba_mod1,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)
```

## Exercise 5

The OR is extremely high for FieldGoals. because the variable is a proportion, so 1 unit increase is not appropriate. We can multiply the variable by 100 to be a percentage

```{r}
nba_cho <- nba_cho |>
  mutate(FieldGoalsPerc = FieldGoals.*100)

nba_mod2 <- glm(Win ~ Home+TeamPoints+FieldGoalsPerc+
                  Assists+Steals+Blocks+TotalRebounds+Turnovers,
                data=nba_cho, family="binomial")
  
modelsummary(nba_mod2,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)
```

## Exercise 6

The odds of winning are 2.46 times higher for home games compared to away games. The relationship between playing at home and winning is statistically significant (p\<0.01, 95% CI: \[1.28,4.82\])

For each additional field goal percent increase, the odds of winning increase by 1.49, which is statistically significant (p\<0.01, 95% CI: \[1.32,1.7\])

etc for steals, total rebounds, and turnovers

## Exercise 7

```{r}
nba_cho_fitted <- factor(ifelse(fitted(nba_mod2)>0.5,1,0),levels=c(0,1),labels=c("Loss","Win"))
confusionMatrix(table(nba_cho_fitted,nba_cho$Win), positive = "Win", mode="everything")

roc(nba_cho$Win, fitted(nba_mod2), print.thres=0.5, print.auc=T,
    legacy.axes=T,plot=T)
```

## Exercise 8

```{r}
nba_mod_reduced <- nba_mod2

nba_cho <- nba_cho |>
  mutate(Opp.FieldGoalsPerc = Opp.FieldGoals.*100)

nba_mod_full <- glm(Win ~ Home+TeamPoints+FieldGoalsPerc+
                  Assists+Steals+Blocks+TotalRebounds+Turnovers+
                    OpponentPoints+Opp.FieldGoalsPerc+Opp.Assists+
                    Opp.Steals+Opp.Blocks+Opp.TotalRebounds+Opp.Turnovers,
                data=nba_cho, family="binomial")

anova(nba_mod_reduced, nba_mod_full, test="Chisq")
```

## Exercise 9

## Bonus

A team always wins when TeamPoints\>OpponentPoints, so there is perfect separation. Removing OpponentPoints eliminates the error.
