---
title: "IDS 702 HW 3"
author: "Your Name Here"
format: pdf
---

**Instructions**: Use this template to complete your assignment. When you click "Render," you should get a PDF document that contains both your answers and code. You must show your work/justify your answers to receive credit. Submit your rendered PDF file on Gradescope. **Remember to render frequently**, as this will help you to catch errors in your code before the last minute.

**Add your name in the Author section in the header**

## Load data

```{r}
library(tidyverse)

college <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/refs/heads/main/colleges.csv")
```

## Exercise 1

a\.

```{r}

```

b\.

```{r}

```

c\.

```{r}

```

| Level | Count (n) | Proportion or % |
|-------|-----------|-----------------|
|       |           |                 |
|       |           |                 |
|       |           |                 |
|       |           |                 |

d\.

```{r}

```

| Level | Count (n) | Proportion or % |
|-------|-----------|-----------------|
|       |           |                 |
|       |           |                 |

e\.

```{r}


```

## Exercise 2

a\.

b\.

c\.

```{r}

```

d\.

```{r}

```

e\.

```{r}

```

f\.

```{r}

```

g\.

h\.

```{r}

```

## Exercise 3

a\.

b\.

c\.

```{r}

```

d\.

```{r}

```

e\.

```{r}

```

f\.

```{r}

```

g\.
