---
title: "Intro to GLMs Group Exercise"
format: pdf
---

Answer the questions below and submit your answers on Gradescope by 3:50PM.

This exercise uses a dataset that contains information on patients and whether or not they were diagnosed with heart disease.

[Source and data dictionary](https://archive.ics.uci.edu/dataset/45/heart+disease)

```{r, message=FALSE}
library(tidyverse)

heart <- read.csv("https://raw.githubusercontent.com/anlane611/datasets/main/heart.csv")
```

1.  Provide summary statistics (N and %) for the outcome variable (`target`)

    ```{r}
    heart |> count(target) |> mutate(perc=(n/sum(n))*100)
    ```

<!-- -->

2.  Calculate summary statistics for those with and without heart disease for the following variables: age, sex, chest pain type, cholesterol. (which summary statistics are appropriate for each variable?)

    ```{r}

    ```

3.  Use the code below to fit a logistic regression model regressing heart disease status on the predictors listed in #3.

```{r}
heart_mod1 <- glm(factor(target) ~ age + factor(sex) + factor(cp) + chol,
                  data=heart,
                  family="binomial")

summary(heart_mod1) #log-odds scale

exp(heart_mod1$coefficients)
```

-   Which variables are statistically significantly associated with heart disease status at the 0.05 level?
-   What are the reference levels for the categorical variables?
-   Write interpretations for the coefficient estimates. Interpret the coefficient estimates in terms of log odds and odds/odds ratios. Are any of these results surprising?
-   Obtain confidence intervals for the coefficient estimates on the log odds and the odds scale.
